---
title: "PB_CDS Analytics"
output: github_document
date: "`r Sys.Date()`"
author: "Khai Nguyen"
description: "This document explores revenue allocation between PB and CDS"
Main.Call.Letter: [PB_CDS]
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## GitHub Documents

This is an R Markdown format used for publishing markdown documents to GitHub. When you click the **Knit** button all R code chunks are run and a markdown file (.md) suitable for publishing to GitHub is generated.

## GitHub Documents

This is an R Markdown format used for publishing markdown documents to GitHub. When you click the **Knit** button all R code chunks are run and a markdown file (.md) suitable for publishing to GitHub is generated.

## Libraries

```{r libraries, message=FALSE, warning=FALSE, include=FALSE}
library(tidyverse)
library(tidyr)
library(readxl)
library(fuzzyjoin)
```

## Load Data
All data are greater than or equal to 1/1/2020
```{r load-data, message=FALSE, warning=FALSE, cache=}
# Load the ZIP code data
fu_ap_gifts <- read.csv("fund_appeal.csv", header = TRUE)
product_gifts <- read.csv("productpurchase.csv", header = TRUE)
```

## Data Exploration and Data Cleaning
Explore the structure and contents of the loaded data frames to understand their format and contents

```{r data-exploration, message=FALSE, cache = TRUE, echo = TRUE}
# fu_ap_gifts exploration
str(fu_ap_gifts)
head(fu_ap_gifts, 5)
tail(fu_ap_gifts, 5)
colSums(is.na(fu_ap_gifts)) # Check for NA values in fu_ap_gifts dataframe

## fu_ap_gifts structure update
### Convert Gift Date from character to Date format
fu_ap_gifts$Gift.Date <- as.Date(fu_ap_gifts$Gift.Date, format = "%m/%d/%Y")

### Convert Gift Amount from character to numeric format (remove $ and commas)
fu_ap_gifts$Gift.Amount <- as.numeric(gsub("[\\$,]",
                                           "",
                                           fu_ap_gifts$Gift.Amount)
)

### Convert Constituent.ID to character format
fu_ap_gifts$Constituent.ID <- as.character(fu_ap_gifts$Constituent.ID)

# product_gifts exploration
str(product_gifts)
head(product_gifts, 5)
tail(product_gifts, 5)
colSums(is.na(product_gifts)) # Check for NA values in product_gifts dataframe

## product_gifts structure update
### Convert Gift Date from character to Date format
product_gifts$Gift.Date <- as.Date(product_gifts$Gift.Date, format = "%m/%d/%Y")

### Convert Gift Amount from character to numeric format (remove $ and commas)
product_gifts$Gift.Amount <- as.numeric(gsub("[\\$,]",
                                              "",
                                              product_gifts$Gift.Amount)
)
### Convert Constituent.ID to character format
product_gifts$Constituent.ID <- as.character(product_gifts$Constituent.ID)
```