---
title: "PB_CDS Analytics"
output: github_document
date: "`r Sys.Date()`"
author: "Khai Nguyen"
description: "This document explores revenue allocation between PB and CDS"
Main.Call.Letter: [PB_CDS]
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## GitHub Documents

This is an R Markdown format used for publishing markdown documents to GitHub. When you click the **Knit** button all R code chunks are run and a markdown file (.md) suitable for publishing to GitHub is generated.

## GitHub Documents

This is an R Markdown format used for publishing markdown documents to GitHub. When you click the **Knit** button all R code chunks are run and a markdown file (.md) suitable for publishing to GitHub is generated.

## Libraries

```{r libraries, message=FALSE, warning=FALSE, include=FALSE}
library(tidyverse)
library(tidyr)
library(readxl)
library(fuzzyjoin)
```

## Load Data
All data are greater than or equal to 1/1/2020
```{r load-data, message=FALSE, warning=FALSE, cache=}
# Load the ZIP code data
fu_ap_gifts <- read.csv("fund_appeal.csv", header = TRUE)
product_gifts <- read.csv("productpurchase.csv", header = TRUE)
```

```{r latest-gift-date}

```