---
title: "PB_CDS Analytics"
output: github_document
date: "`r Sys.Date()`"
author: "Khai Nguyen"
description: "This document explores revenue allocation between PB and CDS"
Main.Call.Letter: [PB_CDS]
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## GitHub Documents

This is an R Markdown format used for publishing markdown documents to GitHub. When you click the **Knit** button all R code chunks are run and a markdown file (.md) suitable for publishing to GitHub is generated.

## GitHub Documents

This is an R Markdown format used for publishing markdown documents to GitHub. When you click the **Knit** button all R code chunks are run and a markdown file (.md) suitable for publishing to GitHub is generated.

## Libraries

```{r libraries, message=FALSE, warning=FALSE, include=FALSE}
library(tidyverse)
library(tidyr)
library(readxl)
library(fuzzyjoin)
```

## Load Data
All data are greater than or equal to 1/1/2020
```{r load-data, message=FALSE, warning=FALSE, cache=}
# Load the ZIP code data
fu_ap_gifts <- read.csv("fund_appeal.csv", header = TRUE)
product_gifts <- read.csv("productpurchase.csv", header = TRUE)
```


## Data Transformation
Transform the data frames to prepare for analysis, such as filtering, selecting relevant columns, and creating new variables

### product_gifts transformation
Categorize the benefits into either listen, read, non_gift, sales_tax, shipping, combo, or other based on benefit.initials.
The benefit.initials variable is created by extracting the first 2 characters of the Benefit variable.
The categorization is based on the productpurchase_categories.csv file
The benefit.category variable is created based on the benefit.initials variable using case_when function.

```{r product_gifts_transformation, message=FALSE, cache = TRUE, echo = TRUE}
# New variable, Fund.Description, with value "Product Purchase"
str(product_gifts)
dim(product_gifts)

product_gifts <- product_gifts %>%
  mutate(Fund.Description = "Product Purchase") %>%
  relocate(Fund.Description, .after = Constituent.ID)

# New variable, benefit.initials, extracting the first 2 characters of Benefit
product_gifts$benefit.initials <- substr(product_gifts$Benefit, 1, 2)
unique(product_gifts$benefit.initials)

# QA benefit.initials to categorize them. Check productpurchase_categories.csv
product_gifts %>%
  filter(product_gifts$benefit.initials == "E-") %>%
  View()

# QA benefit.initials == "E-" to resolve its benefits categorization
product_gifts %>%
  filter(product_gifts$benefit.initials == "E-") %>%
  distinct(Benefit) %>%
  View()

# Replace benefit.initials == "E-" based on the Benefit value
product_gifts <- product_gifts %>%
  mutate(benefit.initials = case_when(
    benefit.initials == "E-" & grepl("CD", Benefit, useBytes = TRUE) ~ "CD",
    benefit.initials == "E-" & grepl("PB", Benefit, useBytes = TRUE) ~ "PB",
    benefit.initials == "E-" & grepl("HB", Benefit, useBytes = TRUE) ~ "HB",
    benefit.initials == "E-" & grepl("MC", Benefit, useBytes = TRUE) ~ "MC",
    benefit.initials == "E-" & grepl("PT", Benefit, useBytes = TRUE) ~ "PT",
    TRUE ~ benefit.initials
  ))

# Categorize records based on benefit.initials
product_gifts <- product_gifts %>%
  mutate(benefit.category = case_when(
    benefit.initials %in% c("Ne", "PB", "Tr", "HB", "SG", "Ca", "PA", "Ma",
                            "BK", "Gr", "WB", "EB", "EP", "PT", "TD", "Fe",
                            "Au", "En", "E ") ~ "read",
    benefit.initials %in% c("MC", "CD", "We", "DV", "Si", "AB", "Mo", "DD",
                            "MP") ~ "listen",
    benefit.initials == "CM" ~ "combo",
    benefit.initials %in% c("Co", "Sp", "Re", "Ra", "Ba", "Ro", "Bo", "Go",
                            "TK") ~ "non_gift",
    benefit.initials == "Sa" ~ "sales_tax",
    benefit.initials %in% c("Sh", "Ex") ~ "shipping",
    TRUE ~ "other")
  )

# QA benefit.category
product_gifts %>%
  distinct(benefit.category) %>%
  View()

# QA data structure
dim(product_gifts)
head(product_gifts, 10)
str(product_gifts)
```

### fu_ap_gifts transformation
Categorize the benefits into either listen, read, combo, 0 normal, or other based on benefit.initials.
The benefit.initials variable is created by extracting the first 2 characters of either Benefit_Premium or Appeal.Description, whichever is not NA.
The categorization is based on the productpurchase_categories.csv file

```{r fu_ap_gifts_transformation, message=FALSE, cache = TRUE, echo = TRUE}
# QA if both Benefit_Premium and Appeal.Description are filled for any record
head(fu_ap_gifts, 1)
fu_ap_gifts %>%
  filter(
    !is.na(Benefit_Premium) & Benefit_Premium != "" &
    !is.na(Appeal.Description) & Appeal.Description != ""
  ) %>%
  count()
# There are 0 records with both features filled

# New variable, benefit.initials, extracting the first 2 characters of Benefit_Premium or Appeal.Description
fu_ap_gifts <- fu_ap_gifts %>%
  filter(!is.na(Benefit_Premium) | !is.na(Appeal.Description)) %>%
  mutate(benefit.initials = case_when(
    !is.na(Benefit_Premium) ~ substr(Benefit_Premium, 1, 2),
    !is.na(Appeal.Description) ~ substr(Appeal.Description, 1, 2)
  ))
str(fu_ap_gifts)

# QA all distinct benefit.initials values
fu_ap_gifts %>%
  distinct(benefit.initials) %>%
  View()

# QA benefit.initials to categorize them. Check productpurchase_categories.csv
fu_ap_gifts %>%
  filter(fu_ap_gifts$benefit.initials == "E-") %>%
  distinct(Benefit_Premium, Appeal.Description) %>%
  View()

# Replace benefit.initials == "E-" based on the Benefit value
fu_ap_gifts <- fu_ap_gifts %>%
  mutate(benefit.initials = case_when(
    benefit.initials == "E-" & grepl("CD", Benefit_Premium, useBytes = TRUE) ~ "CD",
    benefit.initials == "E-" & grepl("PB", Benefit_Premium, useBytes = TRUE) ~ "PB",
    benefit.initials == "E-" & grepl("HB", Benefit_Premium, useBytes = TRUE) ~ "HB",
    benefit.initials == "E-" & grepl("MC", Benefit_Premium, useBytes = TRUE) ~ "MC",
    benefit.initials == "E-" & grepl("PT", Benefit_Premium, useBytes = TRUE) ~ "PT",
    benefit.initials == "E-" & grepl("2024 Wh", Benefit_Premium, useBytes = TRUE) ~ "PB",
    TRUE ~ benefit.initials
  ))

# Categorize records based on benefit.initials
fu_ap_gifts <- fu_ap_gifts %>%
  mutate(benefit.category = case_when(
    benefit.initials %in% c("HB", "PB", "Ne", "Ca", "TD", "WB",
                            "PT", "01", "Ma", "Fe", "Ju", "Au",
                            "Se", "EP", "E ", "En") ~ "read",
    benefit.initials %in% c("CD", "MC", "DV", "DD") ~ "listen",
    benefit.initials == "CM" ~ "combo",
    benefit.initials == "" ~ "0 normal",
    TRUE ~ "other")
  )

# QA benefit.category
fu_ap_gifts %>%
  distinct(benefit.category) %>%
  View()

# New variable, Benefit.Count, counting the number of benefits for each record
fu_ap_gifts <- fu_ap_gifts %>%
  mutate(Benefit.Count = if_else(
    (!is.na(Benefit_Premium) & Benefit_Premium != "") |
      (!is.na(Appeal.Description) & Appeal.Description != ""),
    1, 0
  ))

# New variable, Benefit, consolidating Benefit_Premium and Appeal.Description
fu_ap_gifts <- fu_ap_gifts %>%
  mutate(Benefit = coalesce(
    na_if(Benefit_Premium, ""),
    na_if(Appeal.Description, "")
  ))

# Remove the original Benefit_Premium and Appeal.Description columns
fu_ap_gifts <- fu_ap_gifts %>%
  select(-Benefit_Premium, -Appeal.Description)
str(fu_ap_gifts)
```